\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{ulem}
\newcounter{ProblemNum}
\newcounter{SubProblemNum}[ProblemNum]
\newcommand{\divides}{\bigm|}
\renewcommand{\theProblemNum}{\arabic{ProblemNum}}
\renewcommand{\theSubProblemNum}{\alph{SubProblemNum}}
\newcommand*{\anyproblem}[1]{\newpage\subsection*{#1}}
\newcommand*{\problem}[1]{\stepcounter{ProblemNum} %
\anyproblem{Problem \theProblemNum. \; #1}}
\newcommand*{\soln}[1]{\subsubsection*{#1}}
\newcommand*{\solution}{\soln{Solution}}
\renewcommand*{\part}{\stepcounter{SubProblemNum} %
\soln{Part (\theSubProblemNum)}}


% Document metadata




% Document starts here
\begin{document}
\begin{center}
\begin{Large}
  \textbf{EECS3101 Summer 2024 Assignment 3 }\\
\end{Large}
\begin{large}
	Due: July 31st 23:59
\end{large}
\end{center}
%\vspace{-.5in}

\section*{General Instructions}
Please read the following instructions carefully before starting the exercise. They contain important
information about general exercise expectations, exercise submission instructions,
and reminders of course policies.

\begin{itemize}
\item Your problem set is graded on both correctness and clarity of communication. Solutions
which are technically correct but poorly written will not receive full marks. Please read over
your solutions carefully before submitting them.

\item Each problem set must be completed individually

\item Solutions must be typeset electronically, submitted as a PDF with the correct filename ps1.pdf. Our recommendation goes for using {\LaTeX}  and we recommend \href{https://www.overleaf.com/}{Overleaf} as a tool, but you may feel free to pick your own tool, or generate a PDF using means such as Microsoft Word or other software.


\item Submissions must be made \emph{before} the due date and time on eclass. Late submissions are not accepted.

\end{itemize}



\problem{}
%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%
\textsc{(50 Marks)} {\bf (Greedy Algorithms.)} A group of friends is organizing a bike racing competition. There are $n$ friends and $m$ bikes, but bikes ate not equal, some of them perform better and some of them perform worse. You may think that each bike has a performance factor $p_i, 1\le i\le m$ in the range $(1, 100)$ where $1$ is the worst and $100$ is the best. Each frieng has a greed factor $g_i, 1\le i\le n$ which is the minimum bike performace this firend will be content with. Your goal is to maximize the number of content friends, i.e., friends $i$ assigned a bike $j$ with $g_i\le p_j$. Give a correct greedy algorithm for this problem (15 marks), prove that it finds the optimal solution (20 marks), and give an efficient implementation (15 marks).


%%Write your solution here

\problem{}
\textsc{(50 marks)} {\bf (Greedy Algorithms)}
Describe an efficient algorithm that, given a set $\{x_1, x_2, \dots, x_n\}$ of points on the
real line, determines the smallest set of unit-length closed intervals that contains
all of the given points.  Give a correct greedy algorithm for this problem (15 marks), prove that it finds the optimal solution (20 marks), and give an efficient implementation (15 marks).
%%Write your solution here
\end{document}
